Log or not - Analysis of ECDC forecasts

### 2022-01-31 analysis/Analysis ECDC forecasts.Rmd
add plot with mean observed values across locations

library(scoringutils)
library(purrr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(stringr)
library(here)
library(data.table)
library(kableExtra)
library(performance)

recompute_scores <- FALSE
```

@@

>>>>

library(scoringutils)
library(purrr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(stringr)
library(here)
library(data.table)
library(kableExtra)
library(performance)
library(tidytext)

recompute_scores <- FALSE

# situations with negative values: France 2021-05-22, CZ 2021-08-07, ES 2021-03-06
```

@@

## Mean vs. sd of observed values

```{r}

label_fn <- function(x) {
  ifelse(x < 1000, 
         paste(x), 
         ifelse(x < 1e6, 
                paste0(x / 1000, "k"),
                ifelse(x < 1e9, 
                       paste0(x / 1e6, "m"), 
                       paste0(x / 1e9, "b"))
                )
         )
}

# average number of cases and deaths
hub_data |>
  dplyr::select(location, target_type, target_end_date, true_value) |>
  unique() |>
  group_by(target_type) |>
  summarise(mean_value = mean(true_value), 
            sd_value = sd(true_value), 
            var_value = var(true_value))

# negative binomial
# MASS::glm.nb(true_value ~ target_type, 
#        data = hub_data %>%
#          mutate(true_value = pmax(true_value, 0)), 
#        link = "identity")

# plot with means across locations ---------------------------------------------
label_fn_within <- function(x) {
  x <- gsub(pattern = "___Deaths", "", x)
  x <- gsub(pattern = "___Cases", "", x)
  paste(x)
}

hub_data |>
  dplyr::select(location, target_type, target_end_date, true_value) |>
  unique() |>
  group_by(target_type, location) |>
  summarise(mean_value = mean(true_value), 
            sd_value = sd(true_value), 
            var_value = var(true_value)) |>
  ggplot(aes(y = mean_value, x = reorder_within(location, -mean_value, target_type))) + 
  geom_bar(stat = "identity") + 
  facet_wrap(~ target_type, scales = "free") + 
  theme_minimal() + 
  scale_x_discrete(guide = guide_axis(n.dodge=2), labels = label_fn_within) + 
  labs(y = "Mean observed value", x = "Location")

ggsave("output/figures/HUB-mean-obs-location.png", width = 9, height = 4)

# plot with mean vs. variance --------------------------------------------------
hub_data |>
  dplyr::select(location, target_type, target_end_date, true_value) |>
  unique() |>
  group_by(location, target_type) |>
  summarise(mean_value = mean(true_value), 
            sd_value = sd(true_value), 
            var_value = var(true_value)) |>
  ggplot(aes(y = var_value, x = mean_value)) +
  geom_point() + 
  facet_wrap(~ target_type, scales = "free") + 
  scale_x_continuous(trans = "log10", labels = label_fn) +
  scale_y_continuous(trans = "log10", labels = label_fn) +
  # scale_x_continuous(labels = label_fn) +
  # scale_y_continuous(labels = label_fn) +
  labs(y = "Variance observed value", x = "Mean observed value") + 
  geom_function(fun = function(x) x^2, linetype = "dashed", size = 0.5) + 
  theme_minimal()

ggsave("output/figures/HUB-mean-var-obs.png", width = 7, height = 4)
